---
title: "Sampling"
---

Follow along with the demo to learn how to sample data from a population using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load 1,000 rows of data with the ages of people from 18-80

```{r}
set.seed(1234)

population <- tibble(
  age = sample(18:80, 1000, replace = TRUE)
)
```

### Task 3: Sample 100 rows of data from the dataset

```{r}
sample50 <- population |>
  sample_n(50)
```

### Task 4: Calculate the mean age and standard deviation of the sample

```{r}
sample50 |>
  summarize(
    mean_age = mean(age),
    sd_age = sd(age)
  )
```

### Task 5: Calculate the mean age and standard deviation of the population

```{r}
population |>
  summarize(
    mean_age = mean(age),
    sd_age = sd(age)
  )
```

### Task 6: Compare the sample and population statistics

Compare the means and standard deviations. Are they similar? Go back and re-run the sampling and statistics calculations with a smaller sample size. How do the results change?

The means and standard deviations are very similar. The means are 49.82 for the sample and 49.06 for the population. The standard deviations are 17.88 for the sample and 17.77 for the population. Changing the sample size to 50 slightly changed both values. The mean is now 52.92 and the standard deviation is 17.69. This makes the mean more different from the population but not it is not a significant difference.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
