---
title: "UMD Fees"
---

Follow along with the demo to learn how to calculate percentage change using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load the UMD Fees dataset from <https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv>

```{r}
umd_fees <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv")
```

### Task 3: Calculate the percentage change in each category between Fall 2026 and Fall 2021 using `mutate()`

```{r}
umd_fees_pct <- umd_fees |> mutate(pct_change = (`Fall 2026`- `Fall 2021`)/`Fall 2021` * 100)
```

### Task 4: Describe the results

Write a few sentences about what you found in those percentages. What's the story here? What questions do you have?

There's been a big emphasis on student wellness in the increasing fees from 2021 to 2026. For example, student sustainability has gone up 150% and the student cousneling center was introduced. I wonder where that emphasis came from and the reasoning behind it.

When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
