---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: <https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv>. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")
```

### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

The more important categories are likely the more newsworthy crimes such as homicide. The categories of the data are very specific. For example, assault, breaking and entering and robbery are broken up into different kinds of each crime. This can make summarizing the data more difficult as each specific category being grouped together would make it easier to make calculations for those crimes.

### Task 4: Calculating Averages

Add up the total number of incidents per month for each category using `mutate()`, calling the new column `total_incidents`. Then calculate the average number of incidents per month for each category of crime using the total number of months, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

```{r}
pg_crime <- pg_crime |> mutate(total_incidents = Jan + Feb + Mar + Apr + May + Jun + Jul + Aug + Sep + Oct + Nov + Dec)
pg_crime_averages <- pg_crime |> mutate(average_incidents = total_incidents/12)
```

### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

Categories involving vehicles average the most incidents per month. More deadly crimes such as homicide and sex offense have the lowest average per month.

### Task 6: Testing a Theory

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.

```{r}
summer_crime_averages <- pg_crime_averages |> mutate(
  june_change = (Jun - average_incidents)/average_incidents * 100,
  july_change = (Jul - average_incidents)/average_incidents * 100,
  august_change = (Aug - average_incidents)/average_incidents * 100)
```

### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?

Violent and more serious crimes such as homicide increase in summer months but less serious crimes such as breaking and entering decrease in summer months. This says that the theory's accuracy depends on which specific category of crime you want to look at. If the theory focuses on all crimes then it is likely more inaccurate than accurate.

### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

The majority of Prince George's County crime comes from vehicle-related incidents.

Data from 2024 shows that there were over 5,000 reports of accidents and over 4,000 reports of both stolen vehicles and theft from vehicles. This makes up for the vast majority of crimes that were reported in the county.

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
