---
title: "HW3 WH Salaries"
---

In this assignment, we will be working with a dataset containing information about the salaries of White House staffers as of July 2026.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: <https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2026.csv>. You will create a dataframe called `wh_salaries` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
wh_salaries <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2026.csv")
```

### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, what stands out to you about this data? Are there any salary values that will cause problems for analysis?

This dataframe shows the yearly salaries of employees who work in the White Huse. Something that stands out is the 10 employees who have a yearly salary of \$0, which can also cause problems for analysis as they will significantly any calculations we do such as mean.

### Task 4: Calculate the Mean and Median of the Salaries

Write code to calculate the mean and median of the salaries in the dataset using the `summarize()` function. You don't need to make a new dataframe, just display the results.

```{r}
wh_salaries |> summarise(
  average_salary = mean(SALARY),
  median_salary = median(SALARY)
)
```

### Task 5: Add in the minimum and maximum salaries

Take the code you just wrote and add in the minimum and maximum salaries (the functions for that are `min()` and `max()`). Your results should show four columns: mean_salary, median_salary, min_salary, max_salary.

```{r}
wh_salaries |> summarise(
  average_salary = mean(SALARY),
  median_salary = median(SALARY),
  min_salary = min(SALARY),
  max_salary = max(SALARY)
)
```

### Task 6: Describe the Results

Write a couple of sentences describing the results of your analysis. What do the mean and median salaries tell you about the data? What do the minimum and maximum salaries tell you? Are there any outliers in the data?

The mean and median salaries are very similar, as they only differ by about \$3,000. I was surprised to see this because I thought the 10 \$0 salaries would bring down the mean. Additionally, it was interesting to see that the max salary of \$197,200, which is held by 9 people, are all detailees and three of them are part-time.

### Task 7: Story Idea

Based on the data, what is your best story idea? What is the next step you would take to develop that story idea?

My best story idea is to explore the nine workers who earn \$197,200 and what exactly they do in the White House. Next steps to develop this story would be to research each individual and learn about their specific roles as detailees. Additonally, it could be interesting to compare these nine employees to the 10 employees who do not have a yearly salary.
