---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
library(janitor)
```

Then you'll load the data from the web. The URL for the data is: <https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv>. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
crime <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv") |> clean_names()
```

### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- crime |> mutate(
  violent_rate_2019 = violent2019/pop2019 * 1000,
  property_rate_2019 = property2019/pop2019 * 1000,
  violent_rate_2020 = violent2020/pop2020 * 1000,
  property_rate_2020 = property2020/pop2020 * 1000
)
```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

It is interesting to see that both violent and property crime rates went up in 2020 for some cities. I assumed the COVID-19 pandemic would cause both rates to plummet in the vast majority of cities but that is not that case in this dataset.

### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities_rates |> mutate(
  violent_change = (violent2020 - violent2019)/violent2019 * 100,
  property_change = (property2020 - property2019)/property2019 * 100
)
```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?

The percentage changes are mixed across the data as crime went up in some cities and down in others. It is interesting to see that that the crime changes do not match in a lot of cities. For example, the violent crime rate in Easton went up 79% in 2020 but property crime decreased by 21%. Once again, I am surprised to see crime rates go up in cities despite the COVID-19 pandemic.

### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the *best* story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Violent and property crime rates in some Maryland cities increased in 2020 despite the COVID-19 pandemic.

Crime patterns varied significantly across cities between 2019 and 2020, as some cities experienced increases while others experienced decreases. Some cities such as Frederick saw sharp decreases of around 90% in both violent and property crime in 2020. Other cities such as Gaithersburg saw their violent crime rate increase by 54% and their property crime rate increase by 18%. These numbers show that despite the pandemic keeping people in quarantine, it did not have a consistent affect across Maryland.

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.

### Task 10: Percentiles

10 cities are above the 90th percentile and one city, Cumberland, is at exactly the 90th percentile.

These results do not change my previous story idea as it was about COVID having an inconsistent affect on crime throughout Maryland cities. These results do not necessarily deal with that affect and do not deal with percent change which was a key part of the story.

```{r}
md_cities_percentiles <- md_cities_rates |> mutate(violent_percentile = ntile(violent_rate_2020, 100))

md_cities_rates |>
  summarise(violent_IQR = IQR(violent_rate_2020))
```
