---
title: "HW8_MandMs"
---

In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load the M&Ms dataset from <https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv>

```{r}
mandms <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```

### Task 3: Sample 5 rows of data from the dataset

```{r}
set.seed(42)
sample_5 <- mandms |> sample_n(5)
```

### Task 4: Calculate the mean age and standard deviation of the sample, using either the red or black number of M&Ms

```{r}
sample_5 |>
  summarise(
    mean_red = mean(red),
    sd_red = sd(red)
  )
```

### Task 5: Calculate the mean age and standard deviation of the population, using the same measure as in Task 4

```{r}
mandms |>
  summarise(
    mean_red = mean(red),
    sd_red = sd(red)
  )
```

### Task 6: Compare the sample and population statistics

The two sets of means and standard deviation are somewhat similar, especially considering the small sample size of 5. The sample's mean is 79.6 compared to the population's 72.98 while the sample's standard deviation is 10.06 compared to the population's 16.46.

### Task 7: Describe the results

Write a few sentences about what you found in the data. What does this tell you about the M&Ms dataset and about the act of random sampling?

The high standard deviation of the population data tells me that the amount of red M&Ms for each group is really spread out. This is likely the case for the black M&Ms too as it is totally random how many of each color you get. My finding in the data also tell me that random sampling is the best way to get accurate data of a population. However, the larger the sample the better as the sample size of 5 in this activity led to somewhat high differences in the mean and standard deviation.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
